ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura Gati).}

### 2018-06-28 Latex/Draft_Summer2018/Summer2018.tex
Empirical part - beginning

Regarding the empirical section, the idea is to identify technological shocks which are only specific to the ICT sector in a Structural VAR context.\footnote{An interesting paper which is somehow related to our empirical part is \cite{jafari2012impact}. The authors identify ICT shocks as a potential driver of the Iranian business cycle using a completely different identification strategy and obtaining qualitatively different results.} In order to have a reliable identification procedure our multivariate system needs to embody three key variables: TFP, ICT investment (hereafter ICTI), and relative prices (hereafter RP). Importantly, ICTI is defined as the total expenditure in equipment and computer software meant to be used in production for more than an year. Thus, an increase in ICTI has to be though as an ICT capital deepening. Moreover RP is the ratio between prices in the ICT sector over prices in the overall economy. To verify that we are correctly identifying an ICT technology shock, we firstly expect it to be orthogonal to the productivity of all the other sectors. Since the share of the ICT sector accounts for a negligible part in the whole economy we expect it should have an approximately zero effect on TFP on impact. Moreover, as pointed out by \cite{greenwood1997long} and \cite{fisher2006dynamic}, embodying RP and ICTI is important because we expect that a sectoral technology shock should decrease its relative prices and enhance expenditure in the underlying sector.\footnote{However, as suggested by both \cite{greenwood2000role} and \cite{basu2010sector}, we are aware that conditioning our identifying restrictions only on the direction of RP does not properly measure for technological changes between sectors. This is the main reason why we never impose the direction of RP as an direct identifying condition.} In response to this shock, ICTI rises on impact and remains significant for several quarters. RP persistently and significantly declines for more than two years and TFP, which does not react on impact, rises after few quarters and remains significant and stable for at least 5 years.

@@

In this section we present our main empirical set of results. Our attempt is to properly identify technological shocks which are only specific to the ICT sector in a Structural VAR context and analyze their impact on key macroeconomic variables. At the best of our knowledge, there is only another paper that attempts a similar exercise: \cite{jafari2012impact}. Although also this paper identifies ICT shocks, the authors use Iranian data and both identification strategy and main results are qualitatively different.

@@

In this section we present our main specification where we impose minimal discipline on the identification strategy. It turns out that those set of results presented here are consistent with different robustness checks. Our first-step specification is the following 6-variable VAR

DO NOT READ

\begin{equation}
\begin{pmatrix}
TFP_t \\ 
IT_t \\
GDP_t \\
C_t \\
H_t \\
RP_t \\
\end{pmatrix} = B(L) \begin{pmatrix}
TFP_{t-1} \\ 
IT_{t-1} \\
GDP_{t-1+5|t-1} \\
C_{t-1} \\
H_{t-1} \\
RP_{t-1} \\
\end{pmatrix} + i_t
\end{equation}
where $TFP_t$ is the log-level of Fernald total factor productivity, $H_{t}$ is the level of total number of hours worked per person, $MI_{t+5|t}$ is the level of Michigan Index of business confidence over 5 years, $IT_t$ is the log-level of IT investment, $GDP_t$ is the log-level of real gross domestic product, and $C_t$ is the log-level of real consumption. All the variables refer to the U.S. economy. $B(L)$ is a ($6\times 6$) matrix of lag-operator functions of the same order. In our favorite specification the order of the lag operator functions is two which implies that we regress variables at time $t$ with their own lagged values at $t-1$ and $t-2$. Finally, $i_t$ is a ($6\times 1$) vector of correlated innovations where $\Omega = i'_t i_t$.

In order to derive structural shocks $s_t$ from innovations $i_t$ we use a standard Cholesky identification strategy which imposes 15 zero-impact restrictions. Our identification strategy aims to identify a news shock and a IT productivity shock. In particular, we adhere to the following definitions of our shocks of interest. In line with the news shock literature, a news shock is defined as a signal regarding the future exogenous component of TFP.\footnote{In much of the news shock literature, this definition is not explicit in terms of \emph{exogenous} TFP, because in the absence of an endogenous TFP mechanism, exogenous TFP and total TFP are by definition the same object. Thus, defining a news shock as the shock that moves TFP over the long horizon implicitly amounts to saying that news move exogenous TFP.} The IT productivity shock (also IT shock) is by definition an increase in the sector-specific productivity that should lead to a natural increase in the IT level of investment. %Can we speak here about substitution effect?
The reason why we seek to identify also this second shock is because we want to disentangle the rise in IT investment given by an increase in the sector-specific productivity to the rise in IT investment due to future signals on exogenous TFP. %In other words, in order to size the self-enforcement effect of a news shock we need to disentangle the endogenous component embodied by an IT productivity shock, and, on the other hand, to correctly size the pure effect of an IT productivity shock we need to disentangle from IT its expectational component.

>>>>

Regarding the empirical section, the idea is to identify technological shocks which are only specific to the ICT sector in a Structural VAR context.\footnote{An interesting paper which is somehow related to our empirical part is \cite{jafari2012impact}. The authors identify ICT shocks as a potential driver of the Iranian business cycle using a completely different identification strategy and obtaining qualitatively different results.} In order to have a reliable identification procedure our multivariate system needs to embody three key variables: TFP, ICT investment (hereafter ICTI), and relative prices (hereafter RP). Importantly, ICTI is defined as the total expenditure in equipment and computer software meant to be used in production for more than an year. Thus, an increase in ICTI has to be though as an ICT capital deepening. Moreover RP is the ratio between prices in the ICT sector over prices in the overall economy. To verify that we are correctly identifying an ICT technology shock, we firstly expect it to be orthogonal to the current productivity of all the other sectors. Since the share of the ICT sector accounts for a negligible part in the whole economy we expect it should have an approximately zero effect on TFP on impact. Moreover, as pointed out by \cite{greenwood1997long} and \cite{fisher2006dynamic}, embodying RP and ICTI is important because we expect that a sectoral technology shock should decrease its relative prices and enhance expenditure in the underlying sector.\footnote{However, as suggested by both \cite{greenwood2000role} and \cite{basu2010sector}, we are aware that conditioning our identifying restrictions only on the direction of RP does not properly measure for technological changes between sectors. This is the main reason why we never impose the direction of RP as an direct identifying condition.} In response to this shock, ICTI rises on impact and remains significant for several quarters. RP persistently and significantly declines for more than two years and TFP, which does not react on impact, rises after few quarters and remains significant and stable for at least 5 years.

@@

In this section we present our main empirical set of results. Our attempt is to properly identify technological shocks which are only specific to the ICT sector in a Structural VAR context and analyze their impact on key macroeconomic variables.

@@

In this section we present our main specification where we impose minimal discipline on the identification strategy. It turns out that the set of results presented here are consistent with different robustness checks. Our first-step specification is the following 5-variable VAR 
\begin{equation}
\begin{pmatrix}
TFP_t \\ 
ICT_t \\
GDP_t \\
C_t \\
RP_t \\
\end{pmatrix} = B(L) \begin{pmatrix}
TFP_{t-1} \\ 
ICT_{t-1} \\
GDP_{t-1} \\
C_{t-1} \\
RP_{t-1} \\
\end{pmatrix} + i_t
\end{equation}
where $TFP_t$ is the log-level of Fernald total factor productivity at time $t$, $ICT_{t}$ is the log-level of real information and communication technology investment at time $t$,\footnote{Notice that $ICTI_t$ is defined as the total expenditure at time $t$ in equipment and computer software meant to be used in production for more than an year.} $GDP_t$ is the log-level of real gross domestic product at time $t$, $C_t$ is the log-level of real consumption at time $t$, and $RP_t$ is the log-deviation of ratio between prices of ICT goods and services and the consumer price index (CPI). All the variables refer to the U.S. economy. $B(L)$ is a ($5\times 5$) matrix of lag-operator functions of the same order. Following the Bayesian Information Criterion (BIC), the lag operator functions is one which implies that we regress variables at time $t$ with their own lagged values at $t-1$.\footnote{The Hannan-Quinn Criterion (HQ) suggests to use two lags. Results remains consistent following this second criterion.} Finally, $i_t$ is a ($6\times 1$) vector of correlated innovations where $\Omega = i'_t i_t$.

Using data released on April, 2018 by the Bureau of Economic Analysis (BEA) the real value added of the information sector on real GDP is slightly below $5$\% for the current quarter. Thus, since the share of this sector accounts for a negligible part in the whole economy, as an identification strategy, we assume that an ICT-investment technological change (hereafter ICT shock) is fully orthogonal to current TFP. In addition, in line with theoretical results firstly presented by \cite{greenwood1997long}, we expect that an ICT shock should incentivize sector-specific investment since ICT goods are now easier to produce and cheaper to buy. We transpose this theoretical argument in our empirical VAR by assuming that an ICT shock has a maximal impact effect on ICT investment.
